A Study on Common Bug Fix Patterns in Rust

### 2022-10-04 sections/methodology.tex

\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms. The number of columns would depend on the amount of information we want to encode in the datapoints. The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical. However, in this case we can put a distinction between the aformentioned if statements, as we have a different dimension for nested if statements (\verb+if_if_statement_...+). A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g. a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to JS authors~\cite{foo}). That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset. 
% todo you need to dig up the JS paper ref / Not sure, how to cite, I usually create a bib file, but i'm not sure if i'm allowed to do that for this template.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{figs/essence.png}
    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weight, resulting in the essence of change (Visualized with a circle pack figure)}
\end{figure}

@@

We made two observations about changes. First, we saw that an item (non terminal or BC-related lexemes) that occurs closer to AST leaves tends to be more semantically important. Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

@@

Multiplying the number of occurrences of all observed items in a change by their respective weight results in a vector of numbers. We call this vector, the essence of change. In this vector, the respective value for each item shows its importance in the change. Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}. We have visualized this essence vector using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it.

@@

In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes. Also, we prefer reporting clusters that manifest concrete patterns. An abstract pattern as opposed to a concrete pattern can account for many changes in the programs. For instance, the pattern `Changing a clone of a variable to a borrowing of it in a function argument' is deemed more concrete than `Adding a new statement to a function's body':

>>>>

\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms. The number of columns would depend on the amount of information we want to encode in the datapoints.

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCL}$ as our dimensions, which yields an infeasible $O(n^n)$ dimensions; this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive encodings still holds.

A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g. a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to JS authors~\cite{foo}). That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset. 
% todo you need to dig up the JS paper ref / Not sure, how to cite, I usually create a bib file, but i'm not sure if i'm allowed to do that for this template.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{figs/essence.png}
    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (Visualized with a circle pack figure)}
\end{figure}

@@

We made two observations about changes. First, we saw that an item (non terminal or BC-related lexeme) that occurs closer to AST leaves tends to be more semantically important. Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

@@

Multiplying the number of occurrences of all observed items in a change by their respective weights results in a vector of numbers. We call this vector the \emph{essence of change}. In this vector, the value for each item shows its importance in the change. Figure~\ref{fig:essence} shows the process of computing the essence of the code change that we saw in Figure~\ref{fig:extraction}. We have visualized this essence vector using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it.

@@

In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes. Also, we prefer reporting clusters that manifest concrete patterns. An abstract pattern as opposed to a concrete pattern can account for many changes in the programs. For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more concrete than `Adding a new statement to a function's body':
